\bf Model Visualization Techniques for a Social Network Model

### 2018-05-26 paper-child.Rnw
delete a bunch of unnecessary words. pdf errors on a lot of plots when compiling. no clue what's going on there. i hate this.

<<pkgs>>=
library(tidyverse)
library(RSiena)
library(network)
library(sna)
library(geomnet)
library(GGally)
devtools::install_github("sctyner/netvizinf")
library(netvizinf)
library(RColorBrewer)
library(gridExtra)
library(cowplot)
library(extrafont)
loadfonts(quiet = T)

@@

ggplot(data = toynet$edges) +
  #geom_tile(aes(x = from, y = -to), color = 'grey20', fill = 'black') +
  geom_tile(aes(x = to, y = -from), color = 'grey20', fill = 'black') +
  geom_segment(aes(x = x, xend = xend), y = -0.5, yend=-5.5, colour="grey30", data = data.frame(x = 0:5+.5, xend= 0:5+.5)) +
  geom_segment(aes(y=-y, yend=-yend), x = 0.5, xend=5.5, colour="grey30", data = data.frame(y = -0:5+.5, yend= -0:5+.5)) +
  scale_x_continuous(breaks = 1:5, labels = 1:5, limits = c(0.5,5.5), position = "top") +
  scale_y_continuous(breaks = (-5):(-1), labels = 5:1, limits = c(-5.5,-.5)) +
  ThemeNoNet +
  theme(panel.grid = element_blank(), aspect.ratio = 1,
        plot.title = element_text(size = 20),
        axis.title.x.top = element_text(size = 18),
        axis.title.y.left = element_text(size = 18),
        axis.text.x.top  = element_text(size = 18),
        axis.text.y.left  = element_text(size = 18)) +
  labs(title = "Adjacency Matrix Visualization", x = "\nto (alter)", y = "from (ego)\n")
@
%\hh{XXX Explain each of these visualizations and either comment on the differences in the visualizations (colors) or don't show them. }

@@

\par The underlying SIENA software uses a Robbins-Monro algorithm (see \citet{robbinsmonro}) to estimate the solution of the moment equation
$$ E_{\theta}S = s_{obs}$$
where $\theta = (\boldsymbol{alpha}, \boldsymbol{\beta})$ is the vector of rate and objective function parameters, and $s_{obs}$ is the observed vector of model statistics. The entire algorithm is provided in \citet{SienaAlgs}.

@@

The first way we hope to better understand stochastic actor-oriented models is by viewing the model(s) in the data space. In \citeauthor{modelvis}, they chose define the \textit{data space} as ``the region over which we can reliably make inferences, usually a hypercube containing the data" \citep[p.~206]{modelvis}. But what does this definition mean for network data? The multi-level data structure of networks makes answering this question difficult.
% %For dynamic social networks, there are a few different data ``spaces":
% \begin{enumerate}
% \item the actors and their corresponding covariates,
% \item the edges and their correpsonding variables that describe the relationships between the nodes, and
% \item the time, both the continuous unobserved time and the discrete observed time points, over which the network evolves
% \end{enumerate}
The node, edge, and time data can be visualized together in various ways. %The traditional node-link visualization uses one of many algorithms to layout the actors as points in 2D space, then draws segments connecting the points in 2D if there is an edge between two nodes, and draws nothing otherwise. The time aspect can be visualized by drawing each network observation in time and placing the observed timepoints side-by-side.

Because longitudinal network data consist of three different ``spaces" of data, viewing the model in the data space can depend on which aspect of the \textit{model} and the \textit{data} we are interested in viewing. %The SIENA software includes a method for modelling covariate change in the network
Incorporating data covariates into the network structure allows us to assess whether the ties between nodes are affected by how nodes behave over time. In this instance, we would want to view predictions over time. A SAOM can also model behavior change over time, taking both the node and edge information into account. In this case, a plot of predicted covariate values over time would put the model into the time and node data space. Most likely, however, is that we would want to view all of the data spaces simultaneously.

One tool that can bring the \st{node, edge, and time} data spaces together in this way is the \texttt{R} package \texttt{geomnet} \citep{geomnet}. Different visual features in the node-link diagram can be tied to the underlying node or edge data. The color, size, and shape of the points can be used to represent variables in the node data, while the color, linewidth, and linetype of the lines between points can be used to represent the edge variables. In a social network, node data might be age, gender, and occupation of the person in the network, and edge data might be length of connection between two people, how the people first met (school, work, church, etc.), and how often they interact, and we can view the network at different timepoints side-by-side to see its evolution. Pulling all of this information together with \texttt{geomnet} allows the entire data space to be viewed at once.

@@

To demonstrate, we use \texttt{geomnet} to visualize the connections in the $111^{th}$ United States Senate at two different points: when Hillary Clinton was in the senate, and after she left to become Secretary of State. Clinton was only in the $111^{th}$ senate for 17 days, from January 3, 2009, to January 20, 2009, when she was in the middle of her second term as senator from New York. In that time, she authored two bills and was a cosponsor on 17 other bills. With Clinton included in the node-link diagram, the senate looks much more highly collaborative than it does without her in the diagram. We can compare the number of bills authored throughout the senate by mapping the size of the node to the numer of bills authored by that senator. We also map shape to sex, and keep color mapped to party of the senator. In addition, we can see the strength of the tie by mapping the linewidth of the edge to the $WPC$ value between the two senators. In this single visualization, we have viewed node information (number of bills authored by a senator and the sex and party of that senator), edge information (the direction and strength of ties between two senators), and time (before and after Clinton left the senate).
% ST: I don't like this paragraph
% The visual aspects of the node-link diagram connecting the node, edge, and time data spaces are themselves part of the data space, because they can be viewed in different ways. Another way to connect the node and the edge data, as discussed in Section~\ref{sec:netvizintro}, is through the network's adjaceny matrix. The time dimension can also be incorporated with side-by-side adjacency matrix visualizations.

% below chunk contains some repetitive stuff. Haven't figured out what it's purpose is yet.
<<summnetM1, message=FALSE, fig.cap='On the left, the first wave of observed data that is conditioned on in the model. On the right, the second wave of observed data. In the middle, a summary network from the first model fit to the data. This summary network represents 1,000 simulations of wave 2 using the values from the simple fitted model M1.',results='hide', fig.height=2.5>>=
library(netvizinf)
friendshipData <- array(c(fd2.w1, fd2.w2, fd2.w3), dim = c(16, 16, 3))
drink1 <- s50a[20:35,1]
drink2 <- s50a[20:35,2]
drink3 <- s50a[20:35,3]
friendship <- sienaDependent(friendshipData)
alcohol <- varCovar(cbind(drink1,drink2, drink3))
mydata <- sienaDataCreate(friendship, alcohol)
M1eff <- getEffects(mydata)
M2eff <- includeEffects(M1eff, jumpXTransTrip, interaction1 = "alcohol")
M3eff <- includeEffects(M1eff, nbrDist2twice)
simu2 <- read.csv("data/simulation-1000-M1-M2-M3.csv")
means <- simu2 %>% group_by(Model, parameter) %>% summarize(
  means = mean(estimate)
)
M1parms <- (means %>% filter(Model == "M1"))$means
M2parms <- (means %>% filter(Model == "M2"))$means
M3parms <- (means %>% filter(Model == "M3"))$means
set.seed(4231352)
M1sims1000 <- netvizinf::saom_simulate(dat = mydata,
                            struct = M1eff,
                       parms = as.numeric(M1parms),
                       N = 1000)
M2sims1000 <- netvizinf::saom_simulate(dat = mydata,
                            struct = M2eff,
                       parms = as.numeric(M2parms),
                       N = 1000)
M3sims1000 <- netvizinf::saom_simulate(dat = mydata,
                            struct = M3eff,
                       parms = as.numeric(M3parms),
                       N = 1000)
M1simsdf <- sims_to_df(M1sims1000)
M2simsdf <- sims_to_df(M2sims1000)
M3simsdf <- sims_to_df(M3sims1000)
M1avgW2 <- M1simsdf %>% filter(!is.na(to) & wave == 1) %>%
  group_by(from, to) %>%
  summarise(count = n()) %>%
  mutate(weight = ifelse(from == to, 0, count))
M2avgW2 <- M2simsdf %>% filter(!is.na(to) & wave == 1) %>%
  group_by(from, to) %>%
  summarise(count = n()) %>%
  mutate(weight = ifelse(from == to, 0, count))
M3avgW2 <- M3simsdf %>% filter(!is.na(to) & wave == 1) %>%
  group_by(from, to) %>%
  summarise(count = n()) %>%
  mutate(weight = ifelse(from == to, 0, count))

@@

Another way to view the model in the data space is through simulation from the model. No single network alone simulated from a SAOM is going to look like the data or represent the model, just as no single value simulated from the standard normal distribution will look like a bell-shaped curve. \st{Statisticians would prefer to look at a sample, or at least a summary statistic or two, }so it would therefore better to visualize many simulations together. From a statistician's point of view, a stumbling block with statistical network models generally is the lack of an ``average network" or ``expected network" value. Statisticians frequently rely on averages and expected values in data analyses, but statistical network models, especially those as complex as SAOMs, lack a single, intuitive expected value measure. We could talk about expected values of parameters, but the parameters can be hard to interpret. \st{Expected values of parameters are important, but if they cannot directly tell us anything about our dependent variable (networks in this case), they lose some value to us.} Furthermore, there is no way to talk about the expected value of an observation \st{simulated} from a statistical network model. How then, can we arrive at an ``average" network? We answer this question through visualization. For network data, one way we view an "average" network is through a summary network drawn using the traditional node-link diagram. In Figure~\ref{fig:summnetM1}, we show an \textit{average network} created with 1,000 simulations of the second wave of the network from Model 1. \st{To make this average network, we first simulated 1,000 wave 2 and wave 3 observations of our small friendship example data from model M1, for which parameters had previously been estimated. We then combine the 1,000 instances of wave 2, and count up the number of times each edge appears in a simulation. Then, we combine these 1,000 networks into a single network with edgeweight equal to the proportion of time that edge appears in our 1,000 simulations. This weighted edgelist is the network we draw using the node-link diagram.} An edge is only drawn in the average network if it appears in more than 5\% of the simulations (in at least 51 of the 1000 simulations), with edges that appear more frequently emphasized by thicker linewidths and a darker color, representing the proportion of times appeared in the model simulations. On either side of the average network in Figure~\ref{fig:summnetM1}, we show the actual data, wave 1 on the left, and wave 2 on the right. \st{We can see that the structure of the average network is much more similar to the first wave than to the second wave. However, the simulations are supposed to represent the second wave of data, which is shown on the right in Figure~\ref{fig:summnetM1}. This is an indication that the simple model, M1, is doing a very poor job of capturing the change mechanism from the first to the second wave of observation. The average network can thus be used to help determine model goodness-of-fit. Because the the average network looks more like the first wave than the second wave, we can use the visualization in Figure \ref{fig:summnetM1} as evidence of poor model fit.}  % XXX ST - I was looking at the wrong one. I had wave == 2 instead of wave == 1 in the code. This means I was looking at the wrong simulations, so they really do look more like the first data obs than the second obs. XXX Comparing the three visualizations side-by-side shows us that the summary network of simulations of wave 2 looks more like the true wave 2 than wave 1, which is a good indication that the model is capturing the true data structure. It's not ideal, however, because there appear to be too many reciprocated ties in the summary network, far more than there are in the data. This could indicate that we should remove the reciprocity parameter from the model.

@@

Another potential goodness-of-fit visualization \st{that places the model in the node and edge data space} is a lineup like those proposed in \citet{buja:2009}. A \textit{lineup} ``asks the witness to identify the plot of the real data from among a set of decoys, the null plots, under the veil of ignorance" \citep[p. 4369]{buja:2009}. It can be thought of like a police lineup, where the ``suspect" is in a lineup among several innocent lookalike fillers, and a witness picking the suspect out of the lineup is considered evidence against the suspect. In data and model visualization, the ``suspect" is a plot of the true data, while the ``filler" is composed of several plots of mock data, simulated from a hypothesized model. If the true data stands out among the simulated data, that is taken as evidence of poor model fit, whereas if the true data is difficult to identify among the simulated data, that is taked as evidence of good model fit. \st{An application of the lineup protocal can be found in \citet{hofmann2012graphical}, where the authors examine, for instance, the differences between polar and cartesian coordinates for plotting categorical data, and density plots and box plots for determining distributional differences. They pose questions to experiment participants such as, ``Which plot is most different from the others?" for the first example and, ``In which plot is the blue group furthest to the right?" for the distributional differences. The data visualizations examined in \citet{hofmann2012graphical} are less complex than a node-link network diagram. What questions should we ask for network data visualizations? Asking participants to identify the most different plot may be difficult.  In the network lineup shown in Figure~\ref{fig:lineup}, the second wave of the small friendship network is shown among five simulated networks from model M1 using parameter values estimated from the data. What makes these plots ``different"? It seems possible to argue for any one of the six plots in Figure~\ref{fig:lineup} as most different. So, we guide participants to look at the overall structure of the graphs to determine which has the most and least complex structure. The least complex plot, number six, has no triplets, while the most complex plot, number three, has three triplets, and in fact, plot three is the data. We have found in an experiment where all the node-link diagrams shown are based on simulated data, one observation from an ``alternative" model and the rest from a ``null" model, the triangular shape of the triplets stands out the most to participants. We can use the lineup protocol to help better understand SAOMs because it allows us to view the model in the data space by placing observations the model side-by-side with the data and examining the differences. This can also help us determine the significant structural effects, if any, of the parameters in the model on observations simulated from the model. The triangular shape mentioned above becomes more prevalent when a transitive triplet parameter in included in a model, but it does not always cause triplets to form in the simulated data. This requires more investigation with larger data sets and more complex models, but is promising for the development of additional goodness-of-fit measures for network models.}

@@

\st{There are many possible ways to collect models together. We could look at the same models fit to different data, different models fit to the same data, or because of the nature of SAOMs, we could fit the same models to the same data many times to see how the simulations change. For the SAOMs, we decided that there were four collections that were most important:
\begin{enumerate}
\item the collections resulting from exploring the space of all possible models;
\item the collections we get when varying model settings;
\item the results from fitting the same model form to different data;
\item the results from fitting the same model to the same data many times.
\end{enumerate}

We chose these four collections because they each explore something different about SAOMs. The first takes the many dozens of parameters available to include in a SAOM into account, which easily translates into the second by showing how those many parameters affect the fitted models. Then, we look at how the same model looks when fit to different sets of dynamic network data. Finally, we look at the results from fitting the same model to the same data because the MCMCs and CTMCs that make up a SAOM lead to different parameter estimates every time, so it important to see how the results can vary.}

\subsubsection{Exploring the space of all possible models}

<<sigtesting, fig.show = 'hold', out.width='\\textwidth', fig.height=2, fig.cap="Significant effects for the two data sets, at a significance level of 0.10 or lower as calculated by the Wald-type test available in the SIENA software.">>=
smfr_effs <- read_csv("data/effects_significance_smallFriends.csv")
smfr_effs$inter1[is.na(smfr_effs$inter1)] <- "none"
library(ggrepel)
set.seed(12345678)
smfr_effs %>% filter(type == "eval" & Waldpval <=0.10) %>%
ggplot(aes(x = estimate, y = Waldpval, color = inter1)) +
  #geom_errorbarh(aes(xmin = estimate - se, xmax = estimate + se), color = 'black') +
  geom_vline(xintercept = 0) +
  geom_point(size = 3) +
  #geom_text_repel(aes(label = shortName)) +
  scale_color_brewer(palette = "Paired", name = "Interaction\nVariable") +
  ggtitle("Significant effects at level 0.10 for friendship data") +
  coord_flip() +
  ThemeNoNet
sen_effs <- read_csv("data/congress/senateModelSigEffects.csv")
#sen_effs$Waldpval <- parse_number(sen_effs$Waldpval)
sen_effs$inter1[is.na(sen_effs$inter1)] <- "none"
ggplot(data = sen_effs %>% filter(Waldpval <=0.10), aes(x = estimate, y = Waldpval, color = inter1)) +
  # geom_errorbarh(aes(xmin = estimate - se, xmax = estimate + se), color = 'black') +
  geom_vline(xintercept = 0) +
  geom_point(size = 3) +
  #geom_text_repel(aes(label = shortName)) +
  scale_color_brewer(palette = "Paired", name = "Interaction\nVariable") +
  ggtitle("Significant effects at level 0.10 for senate data") +
  coord_flip() +
  ThemeNoNet
@

\st{The \texttt{RSiena} manual contains over eighty possible effects to include in the model. In order to select parameters to include in the models for our example data, we searched through the possible effects available to model given the data structure to find \textit{significant} effects. We tested for significance using the Wald-type tests built into \texttt{RSiena} for one-at-a-time effects testing. We start with the outdegree and reciprocity measures as the foundation of the models we fit, then add one evaluation effect, fit the model, test the additional effect for significance, and repeat for all possible parameters to add to the model. We performed this procedure for both the small friendship and the senate collaboration data. The results for significant effects are shown in Figure~\ref{fig:sigtesting}. We see in both the friendship data and the senate data results that most of the significant effects have absolute value less than ten. In addition, the $p$-values for the effects from the friendship data are more spread out than the $p$-values for the senate data, which are concentrated at about 0.02 or less. This may suggest that larger data sets tend to result generally in smaller $p$-values, just like a larger sample size results in smaller $p$-values in a $t$-test.}

@@

We have varied model settings already by choosing models M1, M2, M3 to fit to our small friendship network data set. In Section~\ref{sec:samemodsamedat}, we fit these models to the data 1,000 times, and in this section, we explore simulations from these three models given the mean values of from the 1,000 fitted parameter values as the parameters in our models. From each of these three models using the means of parameter estimates as our fixed parameter values, we simulated 1,000 observations from each of the three models. In this process, we condition on the first friendship network observation, and the second and third observations are simulated from the SAOM models with the given parameter values. From these simulations, we first create a visualization that represents an average network.

>>>>

<<pkgs>>=
# if (packageVersion("ggplot2") > "2.2.0"){
#   remove.packages("ggplot2")
#   install.packages("ggplot2")
# }
library(ggplot2)
library(tidyverse)
library(RSiena)
library(network)
library(sna)
library(geomnet)
library(GGally)
#devtools::install_github("sctyner/netvizinf")
library(netvizinf)
library(RColorBrewer)
library(gridExtra)
library(cowplot)
library(extrafont)
loadfonts(quiet = T)

@@

ggplot(data = toynet$edges) +
  #geom_tile(aes(x = from, y = -to), color = 'grey20', fill = 'black') +
  geom_tile(aes(x = to, y = -from), color = 'grey20', fill = 'black') +
  geom_segment(aes(x = x, xend = xend), y = -0.5, yend=-5.5, colour="grey30", data = data.frame(x = 0:5+.5, xend= 0:5+.5)) +
  geom_segment(aes(y=-y, yend=-yend), x = 0.5, xend=5.5, colour="grey30", data = data.frame(y = -0:5+.5, yend= -0:5+.5)) +
  scale_x_continuous(breaks = 1:5, labels = 1:5, limits = c(0.5,5.5), position = "top") +
  scale_y_continuous(breaks = (-5):(-1), labels = 5:1, limits = c(-5.5,-.5)) +
  ThemeNoNet +
  theme(panel.grid = element_blank(), aspect.ratio = 1,
        plot.title = element_text(size = 20),
        axis.title.x = element_text(size = 18),
        axis.title.y = element_text(size = 18),
        axis.text.x  = element_text(size = 18),
        axis.text.y  = element_text(size = 18)) +
  labs(title = "Adjacency Matrix Visualization", x = "\nto (alter)", y = "from (ego)\n")
@
%\hh{XXX Explain each of these visualizations and either comment on the differences in the visualizations (colors) or don't show them. }

@@

\par The underlying SIENA software uses a Robbins-Monro algorithm (see \citet{robbinsmonro}) to estimate the solution of the moment equation
$$ E_{\theta}S = s_{obs}$$
where $\theta = (\boldsymbol{\alpha}, \boldsymbol{\beta})$ is the vector of rate and objective function parameters, and $s_{obs}$ is the observed vector of model statistics. The entire algorithm is provided in \citet{SienaAlgs}.

@@

The first way we hope to better understand stochastic actor-oriented models is by viewing the model(s) in the data space. In \citeauthor{modelvis}, they chose define the \textit{data space} as ``the region over which we can reliably make inferences, usually a hypercube containing the data" \citep[p.~206]{modelvis}. But what does this definition mean for network models? We want to make inferences on parameters in the model, but the multi-level data structure of networks makes the ``region of inference" difficult to define.
% %For dynamic social networks, there are a few different data ``spaces":
% \begin{enumerate}
% \item the actors and their corresponding covariates,
% \item the edges and their correpsonding variables that describe the relationships between the nodes, and
% \item the time, both the continuous unobserved time and the discrete observed time points, over which the network evolves
% \end{enumerate}
%The traditional node-link visualization uses one of many algorithms to layout the actors as points in 2D space, then draws segments connecting the points in 2D if there is an edge between two nodes, and draws nothing otherwise. The time aspect can be visualized by drawing each network observation in time and placing the observed timepoints side-by-side.

The node, edge, and time data can be visualized together in various ways. Because longitudinal network data consist of three different ``spaces" of data, viewing the model in the data space can depend on which aspect of the \textit{model} and the \textit{data} we are interested in viewing. %The SIENA software includes a method for modelling covariate change in the network
Incorporating data covariates into the network structure allows us to assess whether the ties between nodes are affected by how nodes behave over time. In this instance, we would want to view predictions of edge presence over time as node variables change, visualizing all three data spaces at once. In addition, a SAOM can model node variable change over time, taking both the node and edge information into account. In this case, a plot of predicted covariate values over time would put the model into the time and node data space. %It is more intutive to view all of the data spaces simultaneously.

One tool that can bring the node, edge, and time data spaces together is the \texttt{R} package \texttt{geomnet} \citep{geomnet}. Different visual features in the node-link diagram can be tied to the underlying node or edge data. The color, size, and shape of the points can be used to represent variables in the node data, while the color, linewidth, and linetype of the lines between points can be used to represent the edge variables. In a social network, node data might be age, gender, and/or occupation of the person in the network, and edge data might be length of connection between two people, how the people first met (school, work, church, etc.), and/or how often they interact. To view temporal changes, we can view the network at different timepoints side-by-side to see the evolution. Pulling all of this information together with \texttt{geomnet} allows the entire data space to be viewed at once.

@@

To demonstrate, we use \texttt{geomnet} to visualize the connections in the $111^{th}$ United States Senate at two different points: when Hillary Clinton was in the senate, and after she left to become Secretary of State. Clinton was only in the $111^{th}$ senate for 17 days, from January 3, 2009, to January 20, 2009, when she was in the middle of her second term as senator from New York. In that time, she authored two bills and was a cosponsor on 17 other bills. There are many Democratic senators who cosponsored both of her bills, giving their edge a $WPC$ of 1. So with Clinton included in the node-link diagram, the senate looks much more collaborative than it does without her in the diagram. We can compare the Senators' prolificness by mapping the size of the node to the numer of bills authored by the senator. We also map shape to sex, and keep color mapped to party of the senator. In addition, we visualize the strength of the tie by mapping the linewidth of the edge to the $WPC$ value between the two senators. In this single visualization, we have viewed node information (number of bills authored by a senator and the sex and party of that senator), edge information (the direction and strength of ties between two senators), and time (before and after Clinton left the senate). To view the model in the data space, plot simulations from different models over time with applicable node and edge variables using \texttt{geomnet}. 
% ST: I don't like this paragraph
% The visual aspects of the node-link diagram connecting the node, edge, and time data spaces are themselves part of the data space, because they can be viewed in different ways. Another way to connect the node and the edge data, as discussed in Section~\ref{sec:netvizintro}, is through the network's adjaceny matrix. The time dimension can also be incorporated with side-by-side adjacency matrix visualizations.

% below chunk contains some repetitive stuff. Haven't figured out what it's purpose is yet.
<<summnetM1, message=FALSE, fig.cap='On the left, the first wave of observed data that is conditioned on in the model. On the right, the second wave of observed data. In the middle, a summary network from the first model fit to the data. This summary network represents 1,000 simulations of wave 2 using the values from the simple fitted model M1.',results='hide', fig.height=2.5>>=
friendshipData <- array(c(fd2.w1, fd2.w2, fd2.w3), dim = c(16, 16, 3))

drink1 <- drink[20:35,1]
drink2 <- drink[20:35,2]
drink3 <- drink[20:35,3]
friendship <- sienaDependent(friendshipData)
alcohol <- varCovar(cbind(drink1,drink2, drink3))
mydata <- sienaDataCreate(friendship, alcohol)
M1eff <- getEffects(mydata)
M2eff <- includeEffects(M1eff, jumpXTransTrip, interaction1 = "alcohol")
M3eff <- includeEffects(M1eff, nbrDist2twice)
simu2 <- read.csv("data/simulation-1000-M1-M2-M3.csv")
means <- simu2 %>% group_by(Model, parameter) %>% summarize(
  means = mean(estimate)
)
M1parms <- (means %>% filter(Model == "M1"))$means
M2parms <- (means %>% filter(Model == "M2"))$means
M3parms <- (means %>% filter(Model == "M3"))$means
set.seed(4231352)
M1sims1000 <- netvizinf::saom_simulate(dat = mydata,
                            struct = M1eff,
                       parms = as.numeric(M1parms),
                       N = 1000)
M2sims1000 <- netvizinf::saom_simulate(dat = mydata,
                            struct = M2eff,
                       parms = as.numeric(M2parms),
                       N = 1000)
M3sims1000 <- netvizinf::saom_simulate(dat = mydata,
                            struct = M3eff,
                       parms = as.numeric(M3parms),
                       N = 1000)
M1simsdf <- sims_to_df(M1sims1000)
M2simsdf <- sims_to_df(M2sims1000)
M3simsdf <- sims_to_df(M3sims1000)
M1avgW2 <- M1simsdf %>% filter(!is.na(to) & wave == 1) %>%
  group_by(from, to) %>%
  summarise(count = n()) %>%
  mutate(weight = ifelse(from == to, 0, count))
M2avgW2 <- M2simsdf %>% filter(!is.na(to) & wave == 1) %>%
  group_by(from, to) %>%
  summarise(count = n()) %>%
  mutate(weight = ifelse(from == to, 0, count))
M3avgW2 <- M3simsdf %>% filter(!is.na(to) & wave == 1) %>%
  group_by(from, to) %>%
  summarise(count = n()) %>%
  mutate(weight = ifelse(from == to, 0, count))

@@

Another way to view the model in the data space is through simulation from the model. No one network simulated from a SAOM is going to look like the data or represent the model, just as no single value simulated from the standard normal distribution will look like a bell-shaped curve. As statisticians, we prefer to look at a sample, or at least a summary statistic or two, so it would therefore better to visualize many simulated networks at once. A stumbling block with network models generally is the lack of an ``average network" or ``expected network" value. We frequently rely on averages and expected values in data analyses, but network models, especially those as complex as SAOMs, lack a single, intuitive expected value measure. We could talk about expected values of parameters, but the parameters can be hard to interpret. Expected values of parameters can be computed, but if they cannot directly tell us anything about our dependent variable, they lose value. Furthermore, there is no way to talk about the expected value of an observation simulated from a statistical network model, though there are with most other model simulations. How then, can we arrive at an ``average" network? We answer this question through visualization. For network data, one way we view an ``average" network is through a \textit{summary network} drawn using the traditional node-link diagram. In Figure~\ref{fig:summnetM1}, we show an \textit{average network} created with 1,000 simulations of the second wave of the network from Model 1.

To make this average network, we first simulated 1,000 wave 2 and wave 3 observations of our small friendship example data from model M1, for which parameters had previously been estimated. We then combine the 1,000 simulated instances of wave 2, and count up the number of times each edge appears in a network simulation. Then, we combine these 1,000 networks into a single network with edgeweight equal to the proportion of time that edge appears in our 1,000 simulations. This weighted edgelist is the network we draw using the node-link diagram. An edge is only drawn in the average network if it appears in at least 51 of the 1000 simulations, with edges that appear with greater frequency emphasized by thicker linewidths and a darker color. On either side of the average network in Figure~\ref{fig:summnetM1}, we show the actual data, wave 1 on the left, and wave 2 on the right. We can see that the structure of the average network is much more similar to the first wave than to the second wave. However, the simulations are supposed to represent the second wave of data on the right of Firgure~|ref{fig:summnetM1}. This is an indication that the simple model, M1, is doing a very poor job of capturing the change mechanism from the first to the second wave of observation. The average network can thus be used to help determine model goodness-of-fit. Because the the average network looks more like the first wave than the second wave, we can use the visualization in Figure \ref{fig:summnetM1} as evidence of poor model fit.  % XXX ST - I was looking at the wrong one. I had wave == 2 instead of wave == 1 in the code. This means I was looking at the wrong simulations, so they really do look more like the first data obs than the second obs. XXX Comparing the three visualizations side-by-side shows us that the summary network of simulations of wave 2 looks more like the true wave 2 than wave 1, which is a good indication that the model is capturing the true data structure. It's not ideal, however, because there appear to be too many reciprocated ties in the summary network, far more than there are in the data. This could indicate that we should remove the reciprocity parameter from the model.

@@

Another potential goodness-of-fit visualization that places the model in the node and edge data space is a lineup like those proposed in \citet{buja:2009}. A \textit{lineup} ``asks the witness to identify the plot of the real data from among a set of decoys, the null plots, under the veil of ignorance" \citep[p. 4369]{buja:2009}. It can be thought of like a police lineup, where the ``suspect" is in a lineup among several innocent similar looking people, and if the witness picks the suspect out of the lineup, that is evidence the suspect is guilty.

In data and model visualization, the ``suspect" is a plot of the true data, while the ``filler" is composed of several plots of ``mock data", simulated from a hypothesized model. If the true data stands out among the simulated data, that is taken as evidence agains the hypothesised model. If, however, the true data is difficult to identify among the simulated data, that is taken as evidence in favor of the hypothesized model. An application of the lineup protocal can be found in \citet{hofmann2012graphical}, where the authors examine, for instance, the differences between polar and cartesian coordinates for plotting categorical data, and density plots and box plots for determining distributional differences. They pose questions to experiment participants such as, ``Which plot is most different from the others?" or, ``In which plot is the blue group furthest to the right?"

The data visualizations examined in previous applications such as  \citet{hofmann2012graphical} are less complex than a node-link network diagram. What questions should we ask for network data visualizations? The most neutral choice, asking participants to identify the most different plot, may be difficult.  As an example, consider the network lineup shown in Figure~\ref{fig:lineup}. In this small network lineup, the second wave of the small friendship network is shown among five simulated networks from model M1 using parameter values estimated from the data. What exactly makes these plots ``different" from one another?

It seems possible to argue for any one of the six plots in Figure~\ref{fig:lineup} as most different. So, we suggest guiding participants to look at the overall structure of the graphs to determine which has the most or least connected or complex structure. The least complex plot, number six, has no triangles (3-cycles), while the most complex plot, number three, has three 3-cycles. The most complex plot, number three, is in fact the data. We have found that when shown lineups like that of Figure~\ref{fig:lineup}, online survey participants %where all the node-link diagrams shown are based on simulated data, %one observation from an ``alternative" model and the rest from a ``null" model, 
pick up on the triangular shape of the 3-cycles most often.

In a future paper, we will use the lineup protocol to help better understand SAOMs by allowing us to view the model in the data space. Lineups place observations simulated from the model side-by-side with the data and let the viewer examining the differences. This can also help us determine the significant structural effects, if any, of the parameters in the model on observations simulated from the model. The triangular shape mentioned above becomes more prevalent when a transitive triplet parameter ($\beta_3$ in Section~\ref{sec:models}) is included in a model, indicating that it may have a significant effect on network structure. %, but it does not always cause triplets to form in the simulated data. This requires more investigation with larger data sets and more complex models, but is promising for the development of additional goodness-of-fit measures for network models.

@@

The second method of model visualization we use is visualizing collections of models. There are many possible ways to collect SAO models together. We could look at the same model fit to different data, different models fit to the same data, or we could fit the same models to the same data many times to see how the simulations change. For the SAOMs, we decided that there were four collections of models that would give us the most insight:
\begin{enumerate}
\item the collections resulting from exploring the space of all possible models;
\item the collections we get when varying model settings;
\item the results from fitting the same model form to different data;
\item the results from fitting the same model to the same data many times.
\end{enumerate}

We chose these four collections because they each explore something different about SAOMs. The first considers the many dozens of parameters available to include in a SAOM. The second collection show how those many parameters affect the parameter values of the fitted models. Third, looks at how the same model behaves when fit to different sets of dynamic network data. Finally, the results from fitting the same model to the same data may lead to different parameter estimates, so it important to see how the results can vary.

\subsubsection{Exploring the space of all possible models}\label{sec:allpossible}

The \texttt{RSiena} manual contains over eighty possible effects to include in the model. In order to select parameters to include in the models for our example data, we searched through the possible effects available to model given the data structure to find \textit{significant} effects. We tested for significance using the Wald-type tests built into \texttt{RSiena} for one-at-a-time effects testing. We start with the outdegree and reciprocity measures as the foundation of the models we fit, then add one evaluation effect, fit the model, test the additional effect for significance, and repeat for all possible parameters to add to the model. We performed this procedure for both the small friendship and the senate collaboration data. A visualization of the significant effects for the two example data sets are shown in the appendix. %We see in both the friendship data and the senate data results that most of the significant effects have absolute value less than ten. In addition, the $p$-values for the effects from the friendship data are more spread out than the $p$-values for the senate data, which are concentrated at about 0.02 or less. This may suggest that larger data sets tend to result generally in smaller $p$-values, just like a larger sample size results in smaller $p$-values in a $t$-test.

@@

We have varied model settings already by choosing models M1, M2, M3 to fit to our small friendship network data set. In Section~\ref{sec:samemodsamedat}, we fit these models to the data 1,000 times, and in this section, we explore simulations from these three models given the mean values of from the 1,000 fitted parameter values as the parameters in our models. From each of these three models using the means of parameter estimates as our fixed parameter values, we simulated 1,000 observations from each of the three models. In this process, we condition on the first friendship network observation, and the second and third observations are simulated from the SAOM models with the given parameter values. From these simulations, we create visualizations that represents an average network using the same method described in Section~\ref{sec:minds}. These results are provided in the appendix.

%XXXX ST commented out after I realized I had made the same mistake as figure 7. XX This could be an indication that, despite the highly significant nature of the parameter $\beta_3$ according to the $t$-tests in \texttt{RSiena}, the transitive triplet parameter should not be included because it does not result in a structure similar to the data.
